---
title: "hw1"
author: "Ishaan Dey"
date: "2/4/2021"
output:
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, eval = T)
```

# 1. Import

```{r}
library(haven)
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
library(cowplot)
```

```{r}
flights <- read_dta("version10_RA.dta") %>% tibble()
```

# 2. Data Definitions

-- market - the combination of origin and destination airports

-- origin - the origin airport

-- dest - the destination airport

-- year - year when the travel occurred

-- quarter - quarter when the travel occurred

-- carrier - the carrier who transported the passengers (e.g. American
-- AA)

-- nonstopmiles mkt - nonstop distance in miles

-- totalpassengers - passengers transported by the carrier

-- medianmktfare mkt - median fare charged by the carrier

-- meanmktfare mkt - mean fare charged by the carrier

# 3.

```{r}
summary.stats <- function(x,...){
  c(mean=mean(x, ...),
  sd=sd(x, ...),
  min=min(x, ...),
  lower = quantile(x, ..., 1/4),
  median=median(x, ...),
  upper = quantile(x, ..., 3/4),
  max=max(x,...))
}

num.cols <- c("passmeanmktfare_mkt","mktpass_tkcarrier","nonstopmiles_mkt")

flights.summary <- sapply(select(flights, all_of(num.cols)), summary.stats) %>% as.data.frame()
flights.summary 
```

We can see that for `passmeanmktfare_mkt` and `mktpass_tkcarrier`, there
are many outliers that bring up the maximum significantly over the
median. From the histograms, we see that the distribution of
`nonstopmiles_mkt` is fairly even, and confirms lack of outliers.

```{r}
flights %>% 
  select(all_of(num.cols)) %>%
  gather() %>%
  ggplot(aes(value)) + 
  geom_histogram() +
  facet_grid(~key, scales = 'free_x')
```

# 4

We'll first map all the carriers to a group, and check its distribution

```{r}
carrier_groups = list(AA='Legacy',
                      DL='Legacy',
                      UA='Legacy',
                      WN='LCC',
                      B6='LCC')

flights$carrier_group = flights$tkcarrier %>% recode(!!!carrier_groups, .default='Other')
flights$carrier_group %>% table()
```

```{r}

for (g in flights$carrier_group %>% unique()){
  print(paste0('Carrier Group: ',g,'\n'))
  group.stats <- sapply(flights %>% 
           filter(carrier_group == g) %>% 
           select(all_of(num.cols)), 
         summary.stats) %>%
    as.data.frame()
  print(group.stats) 
}
```
